---
title: "pois_test_tutorial"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{pois_test_tutorial}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(pois.exact.test)
```

## 'pois.test'

Simple example on how to use the function `pois.test`:

```{r}
x1 = 100
T1 = 25
result_pois1 = pois.test(x1, T1)

x2 = c(10, 25)
T2 = c(500, 1200)
result_pois2 = pois.test(x2, T2)
```

The "pois.test" function performs an exact test of a simple null hypothesis about the rate parameter in Poisson distribution, or for the ratio between two rate parameters.

Other than the main data parameters, x and T, other parameters that the function takes in include "r", "alternative", and "conf.level". "r = 1" by default, meaning that the hypothesized rate or rate ratio is 1. "conf.level = 0.95" by default, meaning that the confidence level for the returned confidence interval is 95%. 

The following shows the different options for the same data:

```{r}
x = 100
T = 25
result_pois_two_sided = pois.test(x, T)
result_pois_greater = pois.test(x, T, alternative = "greater")
result_pois_less = pois.test(x, T, alternative = "less")
result_pois_90 = pois.test(x, T, conf.level = 0.9)
result_pois_less_90 = pois.test(x, T, alternative = "less", conf.level = 0.9)
result_pois_r_less_90 = pois.test(x, T, r=5, alternative = "less", conf.level = 0.9)
```

This function works with numeric data only, for both x and T. Furthermore, x and T must have compatible length; x must be finite, nonnegative, and integer; T must be nonnegative; r must be a single positive number; conf.level must be a single number between 0 and 1. Otherwise, an error will be thrown, such as in the following:

```{r}
x = 10
y = c(15, 9)
```

```{r}
x = 12.5
```

```{r}
x = -3
```

```{r}
x = "12"
```


To view the number of events (in the first sample if there are two):

```{r}
result_pois_two_sided$statistic
```

To view the corresponding expected count:

```{r}
result_pois1$parameter
```

To view the p-value of the test:

```{r}
result_pois_90$p.value
```

To view a confidence interval for the rate or rate ratio:

```{r}
result_pois_greater$conf.int
```

To view the the estimated rate or rate ratio:

```{r}
result_pois_less$estimate
```

To view the rate or rate ratio under the null, r:

```{r}
result_pois_less_90$null.value
```

To view the alternative hypothesis:

```{r}
result_pois_r_less_90$alternative
```

To view the character string "Exact Poisson test" or "Comparison of Poisson rates" as appropriate:

```{r}
result_pois2$method
```

And finally, to view the names of the data:

```{r}
result_pois_two_sided$data.name
```

correctness check with original `poission.test` function:

```{r}
for (i in 1:5) {
  x = sample(1:1000000, 1)
  T = sample(1000:9999999, 1)
  x2 = sample(1:1000000, 1)
  T2 = sample(1000:9999999, 1)
  alt = sample(c("two.sided", "less", "greater"), 1)
  ci = runif(1, 0, 1)
  print(all.equal(pois.test(x, T, alternative = alt, conf.level = ci), poisson.test(x, T, alternative = alt, conf.level = ci)))
  print(all.equal(pois.test(c(x, x2), c(T, T2), alternative = alt, conf.level = ci), poisson.test(c(x, x2), c(T, T2), alternative = alt, conf.level = ci)))
}
```

bench mark efficiency check with original `poission.test` function:

```{r fig1, fig.height = 3, fig.width = 7}
library(bench)
x = sample(1:1000000, 1)
T = sample(1000:9999999, 1)
x2 = sample(1:1000000, 1)
T2 = sample(1000:9999999, 1)
alt = sample(c("two.sided", "less", "greater"), 1)
ci = runif(1, 0, 1)
mm1 = bench::mark(pois.test(x, T, alternative = alt, conf.level = ci), 
                  poisson.test(x, T, alternative = alt, conf.level = ci))
mm2 = bench::mark(pois.test(c(x, x2), c(T, T2), alternative = alt, conf.level = ci), 
                  poisson.test(c(x, x2), c(T, T2), alternative = alt, conf.level = ci))
print(rbind(mm1, mm2))
```

